Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-19 papers/clusters.tex
Minor edits to intro

Although only hints of small-scale clustering in the abundance space
are visible in \figurename~\ref{fig:all}, exploration of the data indicates that
known clusters do appear in the abundance space as over-densities.
(In general, the collapse of 15 dimensions down to 2-d projections
will hide, smooth, or dilute any structure; there is no guarantee that
a highly featured distribution in 15-d will show features in
\emph{any} 2-d projection, let alone axis-aligned, human-selected projections.)
This encourages us to look for over-densities automatically in the
abundance space and see if anything found that way would be over-dense
in phase space.
The simplest method for clustering points in $D$-dimensional space is
the \emph{k-means} algorithm (see \citealt{bishop} for a pedagogical
introduction and references to the original literature).
Briefly, the k-means algorithm is to find the points closest (in
$D$-space) to each of $K$ centers, and then take the mean of those
points to update the locations of the centers; iterate to convergence.
The output is the locations of the $K$ centers and the assignments of
all points to centers.
This algorithm is fast and performs well in practice in problems of
this nature; also we are not the first to use the k-means algorithm in
abundance space \citep{gratton}.

@@

We chose a few interesting cases from the high-density clusters
in the $K=256$ run and show them in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are dominated by stars in the known clusters
M13, Pal5, and Sagittarius.
The fourth is a halo structure with high velocity dispersion and
possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss the astrophysical implications of these results in
\sectionname~\ref{sec:discussion}.

The abundance-space clusters shown in the \figurename s
might not look dense in abundance space, but they
are very dense in the 15-dimensional space.  Indeed,
\figurename~\ref{fig:M13} is the densest cluster found in the 15-d
space \emph{by far} (see \figurename~\ref{fig:densities}).
The challenge of this work is to find structure in the high
dimensionality space that is not visible in any 2-d projection.
Although we have by no means any model of the 15-d space, we do know
that the structures shown in the figures are very high in 15-d
density.

We have only shown results from the $K=256$ run of k-means.
This run was chosen because its densest abundance-space clusters map
well onto known stellar clusters.
At $K=128$ the densest abundance-space clusters tended to combine
multiple known stellar clusters into single, large abundance-space
groupings.
At $K=512$ k-means tended to split even mono-abundance groups into
smaller sub-groups.
This is a reminder that k-means has been chosen just for simplicity
and reliability here; it is by no means matched to the discovery of
stellar structures.
An important follow-up project is to build a model of abundance space
that captures the features expected from stellar populations.
Even if we were able to tune k-means in some sense ``perfectly'', it
would still break up many stellar clusters, since they often show
multiple stellar populations.
It is also the case that each of the abundance-space clusters we
\emph{do} show in the \figurename s includes both star-cluster members
and some background contamination, and also is missing some true
star-cluster members.

Although we show only three stellar clusters in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, and \ref{fig:Sgr}, many
other clusters are visible among our densest abundance-space k-means
clusters.
These include M15, M92, and M107, among others.
We haven't asked yet whether we detect---as abundance-space
overdensities---all of the stellar clusters we expect to find.
This \documentname\ is simply a demonstration of the chemical-tagging
concept; an investigation of whether we can construct complete
catalogs of stellar clusters is beyond the scope.

>>>>

Although only hints of small-scale clustering in the abundance space
are visible in \figurename~\ref{fig:all}, exploration of the data indicates that
known clusters do appear in the abundance space as over-densities.
In general, the collapse of 15 dimensions down to 2-d projections
will hide, smooth, or dilute any structure; there is no guarantee that
a highly featured distribution in 15-d will show features in
\emph{any} 2-d projection, let alone axis-aligned, human-selected projections.
This encourages us to look for over-densities automatically in the
abundance space and see if anything found that way would be over-dense
in phase space.
The simplest method for clustering points in $D$-dimensional space is
the \emph{k-means} algorithm (see \citealt{bishop} for a pedagogical
introduction and references to the original literature).
Briefly, the k-means algorithm is to find the points closest (in
$D$-space) to each of $K$ centers, and then take the mean of those
points to update the locations of the centers; iterate to convergence.
The output is the locations of the $K$ centers and the assignments of
all points to centers.
This algorithm is fast and performs well in practice in problems of
this nature; also we are not the first to use the k-means algorithm in
abundance space \citep{gratton}.

@@

% ARC TO DO: Make claim of the fourth being M53.
We chose a few interesting cases from the high-density clusters
in the $K=256$ run and show them in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are dominated by stars in the known clusters
M13, Pal5, and Sagittarius.
The fourth is a halo structure with high velocity dispersion and
possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss the astrophysical implications of these results in
\sectionname~\ref{sec:discussion}.

The abundance-space clusters shown in the \figurename s
might not look dense in abundance space, but they
are very dense in the 15-dimensional space.  Indeed,
\figurename~\ref{fig:M13} is the densest cluster found in the 15-d
space \emph{by far} (see \figurename~\ref{fig:densities}).
The challenge of this work is to find structure in the high
dimensionality space that is not obviously visible in any 2-d projection.
Although we have by no means any model of the 15-d space, we do know
that the structures shown in the figures are very high in 15-d
density.

We have only shown results from the $K=256$ run of k-means.
This run was chosen because its densest abundance-space clusters map
well onto known stellar clusters.
At $K=128$ the densest abundance-space clusters tended to combine
multiple known stellar clusters into single, large abundance-space
groupings.
At $K=512$ k-means tended to split even mono-abundance groups into
smaller sub-groups.
This is a reminder that k-means has been chosen just for simplicity
and reliability here; it is by no means matched to the discovery of
stellar structures.
An important follow-up project is to build a model of abundance space
that captures the features expected from stellar populations.
Even if we were able to tune k-means in some sense ``perfectly'', it
would still break up many stellar clusters, since they often show
multiple stellar populations and non-linear correlations between elemental abundances.
It is also the case that each of the abundance-space clusters we
\emph{do} show in the \figurename s includes both star-cluster members
and some background contamination, and also is missing some true
star-cluster members.

Although we show only three stellar clusters in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, and \ref{fig:Sgr}, many
other clusters are visible among our densest abundance-space k-means
clusters.
These include M15, M92, and M107, among others.
We haven't asked yet whether we detect---as abundance-space
overdensities---all of the stellar clusters we expect to find. (Many
clusters within the \apogee\ data set only contain a few plausible 
members, thereby  complicating any inferences we wish to make about 
detection completeness using simple k-means.) 
This \documentname\ is simply a demonstration of the chemical-tagging
concept; an investigation of whether we can construct complete
catalogs of stellar clusters is beyond the scope.

% ARC to DWH: Are the colors of broken glass indicative of anything?
